# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

Load the libraries we will need:

```{r, message=FALSE}
library(ggplot2)
library(zoo)
library(lattice)
```

Unzip and read the data, converting the column to their correct types.

All processing is done in the read.csv

```{r}
unzip("activity.zip")
data <- read.csv("activity.csv",header = TRUE,sep = ",",
                 colClasses = c("numeric","Date","numeric"))
```

## What is mean total number of steps taken per day?

Create a list of the total steps per day and then display the histogram.

```{r}
totalSteps <- aggregate(list(steps = data$steps), by=list(date = data$date), FUN=sum, na.rm=TRUE)
qplot(date, weight = steps, data = totalSteps, geom = "histogram", binwidth=1)
```

The average number of steps per day and the median of the number of steps per day are:

```{r}
mean(totalSteps$steps)
median(totalSteps$steps)
```


## What is the average daily activity pattern?

Create a list that contains the average number of steps per interval

```{r}
dailyActivity <- aggregate(list(steps = data$steps), by=list(interval = data$interval), FUN=mean, na.rm=TRUE)
xyplot(steps ~ interval, data = dailyActivity, type='l', xlab="Interval", ylab="Number of steps")
```

The interval with the maximum number of steps is

```{r}
dailyActivity[which(dailyActivity[,"steps"] == max(dailyActivity$steps)),"interval"]
```

## Imputing missing values

Report on the number of rows with NA

```{r}
sum(is.na(data))
```

Fill in the missing data with the average number of steps for that interval. The new data frame is called fixedData

```{r}
fixedData <- data
fixedData$steps <- na.aggregate(data$steps, by=data$interval, FUN=mean)
```

Create a list of the total steps per day and then display the histogram.

```{r}
totalSteps <- aggregate(list(steps = fixedData$steps), by=list(date = fixedData$date), FUN=sum, na.rm=TRUE)
qplot(date, weight = steps, data = totalSteps, geom = "histogram", binwidth=1)
```

The average number of steps per day and the median of the number of steps per day are:

```{r}
mean(totalSteps$steps)
median(totalSteps$steps)
```

As an effect of our filling in the missing values, the mean and the median are both slightly higher. Also, the mean and the median are now the same.

## Are there differences in activity patterns between weekdays and weekends?

Add a new column with the weekend, weekdays factor

```{r}
fixedData$dayType = as.factor(weekdays(fixedData$date))
levels(fixedData$dayType) <- c("weekday","weekday","weekend","weekend","weekday","weekday","weekday")
```

Create a list of the total steps per day and display the data in a panel plot, ordered by interval to get clean lines

```{r}
dailyActivity <- aggregate(list(steps = fixedData$steps), by=list(interval = fixedData$interval, dayType = fixedData$dayType), FUN=mean, na.rm=TRUE)
xyplot(steps ~ interval | dayType, data = dailyActivity, type='l', layout = c(1,2), xlab="Interval", ylab="Number of steps")
```